---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# estudy2

An implementation of a most commonly used event study methodology,
including both parametric and nonparametric tests. It contains variety
aspects of the rate of return estimation (the core calculation is done in
C++), as well as three classical market models: mean adjusted returns,
market adjusted returns and single-index market models. There are 6
parametric and 6 nonparametric tests provided, which examine
cross-sectional daily abnormal return (see the documentation of the
functions for more information). Furthermore, tests for the cumulative
abnormal returns are included.

## Instalation
The package is not submitted to CRAN, and is located in a private repository.
If one has an access to the github repo, it is possible to install the last 
development version:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("irudnyts/estudy2")
```
