---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# estudy2

[![Travis-CI Build Status](https://travis-ci.com/irudnyts/estudy2.svg?token=Spwopnmy37EZMsL4nDza&branch=master)](https://travis-ci.com/irudnyts/estudy2.svg?token=Spwopnmy37EZMsL4nDza&branch=master)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/irudnyts/estudy2?branch=master&svg=true)](https://ci.appveyor.com/project/irudnyts/estudy2)
[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/estudy2)](https://cran.r-project.org/package=estudy2)

## Overview

An implementation of a most commonly used event study methodology,
including both parametric and nonparametric tests. It contains variety
aspects of the rate of return estimation (the core calculation is done in
C++), as well as three classical market models: mean adjusted returns,
market adjusted returns and single-index market models. There are 6
parametric and 6 nonparametric tests provided, which examine
cross-sectional daily abnormal return (see the documentation of the
functions for more information). Furthermore, tests for the cumulative
abnormal returns are included.

## Installation

To install a current stable release from CRAN use: 

```{r, eval=FALSE}
install.packages("estudy2")
```


To install the development version of estudy2 use:

```{r, eval=FALSE}
# install.packages("devtools")
# library("devtools")
devtools::install_github("irudnyts/estudy2")
```
